### 2024-05-01 content/chapter_5.tex
:triangular_flag_on_post: Viva

\begin{section}{Introduction}
	% The quest for reliable, high-speed, and ubiquitous wireless connectivity has been long-standing since Marconi's illuminating radio in 1895.
	% Great successes have been made at transmitter and receiver sides over the past century, and the society is unprecedentedly close to the Shannon limit \cite{Shannon1948}.
	Today we are witnessing a paradigm shift from connectivity to intelligence, where the wireless environment is no longer a chaotic medium but a conscious agent that serves on demand.
	This is empowered by the recent advances in \gls{ris}, a real-time programmable metasurface of numerous non-resonant sub-wavelength scattering elements.
	It can manipulate the amplitude, phase, frequency, and polarization of the scattered waves \cite{Basar2019} with a higher energy efficiency, lower cost, lighter footprint, and greater scalability than relays.
	Using \gls{ris} for {passive beamforming} has attracted significant interest in wireless communication \cite{Wu2019,Wu2020c,Yang2020,Zheng2021}, backscatter \cite{Jia2020,Liang2022}, sensing \cite{Liu2022a,Hua2023}, and power transfer literature \cite{Wu2021d,Feng2022,Zhao2022}, reporting a second-order array gain and fourth-order power scaling law (with proper waveform).
	On the other hand, \gls{ris} also enables {backscatter modulation} by dynamically switching between different patterns, as already investigated \cite{Karasik2020,Basar2020,Zhao2022a} and prototyped \cite{Tang2019b,Dai2020a}.
	Despite fruitful outcomes, one critical unanswered question is the {channel shaping} capability: \emph{To what extent can a passive \gls{ris} reshape the wireless channel?}

	The answer depends heavily on the hardware architecture and scattering model.
	In diagonal \gls{ris}, each scattering element is tuned by a dedicated impedance and acts as an \emph{individual} phase shifter \cite{Wu2020}.
	The concept is generalized to \gls{bd}-\gls{ris} \cite{Shen2020a,Li2023b} which physically connects adjacent elements using passive components.
	This allows \emph{cooperative} scattering --- wave impinging on one element can propagate within the circuit and depart partially from any element in the same group.
	\gls{bd}-\gls{ris} can thus control both amplitude and phase of the reflected wave, generalizing the scattering matrix from diagonal with unit-magnitude entries \eqref{eq:diagonal_scattering_matrix} to block diagonal with unitary blocks \eqref{eq:bd_scattering_matrix}.
	Its benefit has been recently shown in receive power maximization \cite{Nerini2023,Santamaria2023,Fang2023,Nerini2024}, transmit power minimization \cite{Zhou2023}, and rate maximization \cite{Zhou2023,Nerini2024,Li2023d,Bartoli2023,Li2023c}.
	Practical issues such as channel estimation \cite{Li2023e} and mutual coupling \cite{Li2023f} have also been investigated.
	Therefore, \gls{bd}-\gls{ris} is envisioned as the next generation channel shaper with stronger signal processing flexibility \cite{Li2023g}.

	% Attempts to characterize the channel shaping capability can be classified into \emph{singular value centric} and \emph{power centric} methods.
	% Channel shaping is different from passive beamforming as it seeks to modify the inherent properties of the channel itself, allowing one to decouple \gls{ris} and transceiver design.
	Channel shaping is different from passive beamforming discussed in Chapters \ref{ch:ris_aided_swipt} and \ref{ch:riscatter} because \gls{ris} is exploited as a stand-alone device to modify the inherent properties of the wireless channel.
	This allows one to decouple the \gls{ris}-transceiver design and explore the fundamental limits of wave manipulation.
	Relevant design criteria can be roughly classified into two categories:
	\begin{itemize}
		\item \emph{Singular value centric:} Focus on the singular values of the equivalent channel matrix, which are closely related to the channel capacity and the diversity-multiplexing trade-off. Relevant designs are more complicated since they involve the entire channel matrix.
		\item \emph{Power centric:} Focus on the second-order characteristics of the channel matrix, which are useful in \gls{miso} but less informative in \gls{mimo}. Relevant designs are more practical and straightforward.
	\end{itemize}
	The former has been studied in terms of minimum singular value \cite{ElMossallamy2021}, effective rank \cite{ElMossallamy2021,Meng2023}, condition number \cite{Zheng2022,Huang2023} in \gls{pc}, and degree of freedom \cite{Bafghi2022,Zheng2023,Chae2023} in \gls{ic}.
	The latter has been studied in terms of channel power \cite{Wu2019,Nerini2023,Santamaria2023} in \gls{pc} and leakage interference \cite{Santamaria2023a} in \gls{ic}.
	Although the results are insightful, the big picture is far from complete since there exists neither a systematic channel shaping study nor a universal \gls{ris} optimization framework.
	To the best of our knowledge, the most general channel shaping scenario (\gls{bd}-\gls{ris}-aided \gls{mimo}) has not been studied, and the great potential of \gls{bd}-\gls{ris} has not been fully exploited or understood.
	This chapter aims for a comprehensive answer to the channel shaping question through theoretical analysis and numerical optimization.
	The contributions are summarized below.

	First, we exploit the Riemannian geometry of the Stiefel manifold to design a novel geodesic\footnote{A geodesic refers to the shortest path between two points in a Riemannian manifold.} \gls{rcg} algorithm for \gls{bd}-\gls{ris} optimization.
	This method modified from \cite{Abrudan2008,Abrudan2009} not only features lower complexity and faster convergence than general non-geodesic approach \cite{Absil2009,Pan2022d}, but also fits arbitrary group size and any smooth \gls{ris} optimization problem.
	Specifically, block-wise rotational update is performed along the geodesics of the Stiefel manifold, which are expressed compactly by the exponential map \cite{Edelman1998}.
	This is the first work to tailor an efficient optimization framework for asymmetric \gls{bd}-\gls{ris}.

	Second, we quantify the capability of a \gls{bd}-\gls{ris} to redistribute the singular values of \gls{mimo}-\gls{pc}.
	The \emph{Pareto frontiers} are characterized by optimizing the {weighted sum of singular values}, where the weights can be positive, zero, or negative.
	This problem is solved by the proposed geodesic \gls{rcg} algorithm.
	The resulting achievable singular value region generalizes most relevant metrics and provides an intuitive channel shaping benchmark.
	We also demystify the \gls{bd}-\gls{ris} gain from off-diagonal entries and derive some analytical singular value bounds for rank-deficient \gls{mimo} and fully-connected (a.k.a. unitary) \gls{ris}.
	This is the first work to comprehensively answer the channel shaping question and fully unveil the potential or \gls{bd}-\gls{ris}.

	Third, we tackle \gls{bd}-\gls{ris}-aided \gls{mimo}-\gls{pc} rate maximization with two solutions: a local-optimal approach through \gls{ao} and a low-complexity approach over channel shaping.
	The former updates active and passive beamforming by eigenmode transmission and \gls{rcg} algorithm, respectively.
	The latter suboptimally decouples both blocks, recasts the shaping problem as channel power maximization, and solves it iteratively in closed form.
	Interestingly, the gap in between vanishes as \gls{bd}-\gls{ris} evolves from diagonal to unitary.
	It suggests channel shaping decouples the \gls{ris}-transceiver design and offers a promising low-complexity solution.
	This is the first work to discuss the implications of channel shaping in joint active-passive beamforming optimization.

	Fourth, we extend the aforementioned approaches to \gls{mimo}-\gls{ic} where the \gls{bd}-\gls{ris} is used for leakage interference minimization and \gls{wsr} maximization.
	In the former case, we update the receive combiner, transmit precoder, and scattering matrix iteratively in closed forms with optimality proof.
	In the latter case, we alternatively updates the transmit precoder by bisection and the scattering matrix by \gls{rcg} algorithm.
	We also compare the impact of \gls{ris} on \gls{pc} and \gls{ic}.
	% provide a comprehensive comparison between the two approaches and discuss the implications of channel shaping in \gls{ic}.

@@

\begin{section}{\glsfmtshort{mimo}-\glsfmtshort{pc}}
	\begin{subsection}{System Model}
		Consider a \gls{bd}-\gls{ris}-aided point-to-point \gls{mimo} system with $N_\mathrm{T}$, $N_\mathrm{S}$, $N_\mathrm{R}$ transmit, scatter, and receive antennas, respectively.
		This configuration is denoted as $N_\mathrm{T} \times N_\mathrm{S} \times N_\mathrm{R}$ in the following context.
		The \gls{bd}-\gls{ris} can be modeled as an $N_\mathrm{S}$-port network \cite{Ivrlac2010} that divides into $G$ individual groups, each containing $L \triangleq N_\mathrm{S} / G$ elements interconnected by real-time reconfigurable components \cite{Shen2020a}.
		To simplify the analysis, we assume a lossless asymmetric network\footnote{While symmetric \gls{bd}-\gls{ris} (involving capacitors and inductors) is often considered in the literature \cite{Shen2020a,Nerini2023,Santamaria2023,Fang2023,Nerini2024,Zhou2023,Li2023d,Bartoli2023}, asymmetric \gls{bd}-\gls{ris} can be built over reconfigurable asymmetric passive components (e.g., ring hybrids and branch-line hybrids) \cite{Ahn2006}.} without mutual coupling between scatter elements, as previously considered in \cite{Li2023b,Li2023c,Bartoli2023}.
		The overall scattering matrix of the \gls{bd}-\gls{ris} is
		\begin{equation}
			\mathbf{\Theta} = \mathrm{diag}(\mathbf{\Theta}_1,\ldots,\mathbf{\Theta}_G),
		\end{equation}
		where $\mathbf{\Theta}_g \in \mathbb{U}^{L \times L}$ is a unitary matrix that describes the response of group $g \in \mathcal{G} \triangleq \{1, \ldots, G\}$.
		That is, $\mathbf{\Theta}_g^\mathsf{H} \mathbf{\Theta}_g = \mathbf{I}$.
		Note that diagonal (single-connected) and unitary (fully-connected) \gls{ris} can be regarded as its extreme cases with group size $L=1$ and $L=N_\mathrm{S}$, respectively.
		Some potential physical architectures of \gls{bd}-\gls{ris} are illustrated in \cite[Fig. 3]{Shen2020a}, \cite[Fig. 5]{Li2023c}, and \cite[Fig. 2]{Nerini2024}, where the radiation pattern and circuit topology have impacts on the scattering matrix.

		Let $\mathbf{H}_\mathrm{D} \in \mathbb{C}^{N_\mathrm{R} \times N_\mathrm{T}}$, $\mathbf{H}_\mathrm{B} \in \mathbb{C}^{N_\mathrm{R} \times N_\mathrm{S}}$, $\mathbf{H}_\mathrm{F} \in \mathbb{C}^{N_\mathrm{S} \times N_\mathrm{T}}$ denote the direct (transmitter-receiver), backward (\gls{ris}-receiver), and forward (transmitter-\gls{ris}) channels, respectively.
		The equivalent channel is a function of the scattering matrix
		\begin{equation}
			\mathbf{H} = \mathbf{H}_\mathrm{D} + \mathbf{H}_\mathrm{B} \mathbf{\Theta} \mathbf{H}_\mathrm{F} = \mathbf{H}_\mathrm{D} + \sum_g \mathbf{H}_{\mathrm{B},g} \mathbf{\Theta}_g \mathbf{H}_{\mathrm{F},g},
			\label{eq:channel_equivalent}
		\end{equation}
		where $\mathbf{H}_{\mathrm{B},g} \in \mathbb{C}^{N_\mathrm{R} \times L}$ and $\mathbf{H}_{\mathrm{F},g} \in \mathbb{C}^{L \times N_\mathrm{T}}$ are the backward and forward subchannels for \gls{ris} group $g$, corresponding to the $(g-1)L$ to $gL$ columns of $\mathbf{H}_\mathrm{B}$ and rows of $\mathbf{H}_\mathrm{F}$, respectively.
		% That is, $\mathbf{H}_{\mathrm{B},g}$ (resp. $\mathbf{H}_{\mathrm{F},g}$) corresponds to the $(g-1)L$ to $gL$ columns (resp. rows) of $\mathbf{H}_\mathrm{B}$ (resp. $\mathbf{H}_\mathrm{F}$).
		% , corresponding to the $(g-1)L$ to $gL$ columns of $\mathbf{H}_\mathrm{B}$ and rows of $\mathbf{H}_\mathrm{F}$.
		% $\mathbf{H}_{\mathrm{B},g}$ (resp. $\mathbf{H}_{\mathrm{F},g}$) corresponds to the $(g-1)L$ to $gL$ columns of $\mathbf{H}_\mathrm{B}$
		Since unitary matrices constitute an algebraic group with respect to multiplication, the scattering matrix of group $g$ can be decomposed as
		\begin{equation}
			\mathbf{\Theta}_g = \mathbf{L}_g \mathbf{R}_g^\mathsf{H},
			\label{eq:unitary_decomposition}
		\end{equation}
		where $\mathbf{L}_g, \mathbf{R}_g \in \mathbb{U}^{L \times L}$ are two unitary factor matrices.
		Let $\mathbf{H}_{\mathrm{B},g} = \mathbf{U}_{\mathrm{B},g} \mathbf{\Sigma}_{\mathrm{B},g} \mathbf{V}_{\mathrm{B},g}^\mathsf{H}$ and $\mathbf{H}_{\mathrm{F},g} = \mathbf{U}_{\mathrm{F},g} \mathbf{\Sigma}_{\mathrm{F},g} \mathbf{V}_{\mathrm{F},g}^\mathsf{H}$ be the compact \gls{svd} of the backward and forward channels, respectively.
		The equivalent channel can thus be rewritten as
		\begin{equation}
			\mathbf{H} = \overbrace{\mathbf{H}_\mathrm{D} + \sum_g \mathbf{U}_{\mathrm{B},g} \mathbf{\Sigma}_{\mathrm{B},g} \underbrace{\mathbf{V}_{\mathrm{B},g}^\mathsf{H} \mathbf{L}_g \mathbf{R}_g^\mathsf{H} \mathbf{U}_{\mathrm{F},g}}_\text{backward-forward} \mathbf{\Sigma}_{\mathrm{F},g} \mathbf{V}_{\mathrm{F},g}^\mathsf{H}}^\text{direct-indirect}.
			\label{eq:channel_equivalent_svd}
		\end{equation}

>>>>

\begin{section}{Introduction}
	Today we are witnessing a paradigm shift from connectivity to intelligence, where the wireless environment is no longer a chaotic medium but a conscious agent that can serve on demand.
	This is empowered by the recent advances in \gls{ris}, a programmable metasurface that recycles and redistributes ambient electromagnetic waves for improved wireless performance.
	A typical \gls{ris} consists of numerous low-power sub-wavelength non-resonant scattering elements, whose response can be engineered in real-time to manipulate the amplitude, phase, frequency, and polarization of the scattered waves \cite{Basar2019}.
	% It not only experiences negligible noise and supports full-duplex transmissions, but also features better flexibility than reflectarrays, lighter footprint than various relays, and greater scalability than conventional multi-antenna techniques.
	It enables low-noise, full-duplex operation, and also features better flexibility than reflectarrays, lighter footprint than various relays, and greater scalability than conventional multi-antenna techniques.
	The most popular \gls{ris} research direction is \emph{joint beamforming} design with transceivers for a specific performance measure, which has attracted significant attention in wireless communication \cite{Wu2019,Guo2020,Liu2022}, sensing \cite{He2022,Luo2022,Hua2023}, and power transfer literature \cite{Wu2020a,Feng2022,Zhao2022}.
	While passive beamforming suffers severe attenuation from double fading, it usually offers a squared asymptotic behavior than active beamforming (e.g., second-order array gain and fourth-order harvested power \cite{Zhao2022}).
	On the other hand, \gls{ris} can also be used for \emph{backscatter modulation} by periodically switching its reflection pattern within channel coherence time.
	This creates a free-ride message stream (similar to index modulation \cite{Basar2017}) with dual benefits: integrating with the legacy transmitter for enhanced channel capacity \cite{Karasik2020,Basar2020,Ye2022} or serving as a dedicated source for low-power uplink communication \cite{Liang2020,Zhao2024,Yang2024}.
	Different from above directions, \emph{channel shaping} exploits the \gls{ris} as a stand-alone device to modify the inherent properties of the propagation environment.
	This provides a ubiquitous wave scattering benchmark for different wireless applications and helps to decouple the \gls{ris}-transceiver design.
	Relevant shaping metrics can be classified into two categories:
	\begin{itemize}
		\item \emph{Singular value centric:} Closely related to the performance measures (e.g., achievable rate and harvested power \cite{Shen2021}) but sensitive to numerical perturbations. The impact of \gls{ris} has been studied in terms of minimum singular value \cite{ElMossallamy2021}, effective rank \cite{ElMossallamy2021,Meng2023}, condition number \cite{Zheng2022,Huang2023}, and degree of freedom \cite{Bafghi2022,Zheng2023,Chae2023}.
		\item \emph{Power centric:} The second-order statistics are less informative in \gls{mimo} but easier to analyze and optimize. The impact of \gls{ris} has been studied in terms of channel power gain \cite{Wu2019,Shen2020a,Nerini2023,Nerini2024,Santamaria2023} and leakage interference \cite{Santamaria2023a}.
	\end{itemize}

	While those works offer initial glimpses into the channel shaping potential, one critical question remains unaddressed: \emph{To what extent can a passive \gls{ris} reshape the \gls{mimo} channel in terms of singular values?}
	% \begin{enumerate}
	% 	\item What is the entire singular value region of a \gls{mimo} channel with \gls{ris}?
	% \end{enumerate}
	% they neither discuss the limitation
	% depict the entire singular value region nor provide a universal optimization framework.
	% It motivates the ultimate question: \emph{To what extent can a passive \gls{ris} reshape \gls{mimo} channels in terms of singular values?}
	% Despite the great potentials and fruitful outcomes, one critical unanswered question is the {channel shaping} capability: \emph{To what extent can a passive \gls{ris} reshape \gls{mimo} channels?}
	The answer depends heavily on the scattering model and hardware architecture.
	Most relevant works \cite{ElMossallamy2021,Meng2023,Zheng2022,Huang2023,Bafghi2022,Zheng2023,Chae2023,Wu2019,Santamaria2023a} assumed that each \gls{ris} element is tuned by a dedicated impedance and acts as an \emph{individual} scatterer \cite{Wu2020}.
	This ideally translates to a scattering matrix with unit-magnitude entries on the main diagonal and zeros elsewhere, applying merely a phase shift to the incoming signal.
	The concept was soon generalized to \gls{bd}-\gls{ris} \cite{Shen2020a} that physically groups adjacent elements using passive reconfigurable components.\footnote{Those components can be either symmetric (e.g., capacitors and inductors) or asymmetric (e.g., ring hybrids and branch-line hybrids) \cite{Ahn2006}, resulting in symmetric and asymmetric scattering matrices, respectively.}
	This allows \emph{cooperative} scattering --- wave impinging on one element can propagate within the circuit and depart partially from any element in the same group.
	It can thus redistribute both amplitude and phase of the scattered wave with zero power loss, generalizing the scattering matrix to block-diagonal with unitary blocks.
	Such a powerful model can be realized at reduced hardware cost using tree- and forest-connected architectures inspired by graph theory \cite{Nerini2024}.
	\gls{bd}-\gls{ris} can also function in hybrid transmitting-and-reflecting mode \cite{Li2023b} and multi-sector mode \cite{Li2023c} to provide full-space coverage and multi-user support.
	Many practical design challenges have been addressed including channel estimation \cite{Li2024}, mutual coupling \cite{Li2023f}, and wideband modelling \cite{Li2024a}.
	Its beamforming superiority has been studied extensively in \gls{siso} and \gls{miso} systems, where the problems were cast as single-user \gls{snr} maximization \cite{Shen2020a,Nerini2023,Nerini2024,Santamaria2023} and multi-user \gls{wsr} maximization \cite{Fang2023,Zhou2023,Li2023c,Soleymani2024}.
	However, the interplay between \gls{bd}-\gls{ris} and \gls{mimo} systems is still in the infancy stage.
	The authors of \cite{Bartoli2023} investigated the rate-optimal joint beamforming design for a specific \gls{bd}-\gls{ris}-aided \gls{mimo} system with blocked direct link and unitary (a.k.a. fully-connected) \gls{ris}.
	Similar constraints were also adopted in \cite{Mishra2024}, which introduces a transmitter-side \gls{bd}-\gls{ris} to massive \gls{mimo} that exploits statistical \gls{csi} for improved spectral efficiency.
	Received power maximization problem was studied over continuous-valued \cite{Nerini2023} and discrete-valued \cite{Nerini2023b} \gls{bd}-\gls{ris}, but the proposed single-stream transceiver is rate-suboptimal and the passive beamforming problem is equivalent to \gls{siso}.
	A practical frequency-dependent \gls{bd}-\gls{ris} model has been recently proposed for multi-band multi-cell \gls{mimo} networks to facilitate practical deployments \cite{Sena2024}.
	Although the results are promising, the channel shaping capability from off-diagonal entries deserves further investigation, and there lacks a universal yet efficient \gls{bd}-\gls{ris} design framework.
	Furthermore, no previous work has attempted to characterize the singular value region of a \gls{mimo} channel manipulated by any type of \gls{ris}, and the \gls{bd}-\gls{ris}-aided \gls{mimo} rate maximization problem remains unsolved.
	This paper aims for a comprehensive answer to the channel shaping question through theoretical analysis and numerical optimization.
	The contributions are summarized below.

	% 1. need a general study on BD-RIS x MIMO and explain the potential from off-diagonal entries
	% 2. need a general optimization framework for BD-RIS
	% 3. need a comprehensive answer to the channel shaping question
	% 4. need a low-complexity solution from shaping

	First, we discuss the shaping potential from off-diagonal entries of \gls{bd}-\gls{ris} in terms of subchannel rearrangement and subspace alignment.
	Subchannel rearrangement enables a flexible in-group permutation and combination of backward and forward channels by their strength.
	This unique feature of \gls{bd}-\gls{ris} provides a higher shaping freedom and exploits spatial diversity more effectively than diagonal \gls{ris}.
	On the other hand, subspace alignment in \gls{mimo} generalizes phase matching in \gls{siso} and \gls{miso} to the high-dimensional singular vector space.
	The scattering matrix simultaneously affects the multiplicative intra-group alignment and additive inter-group alignment.
	However, both subspaces cannot be perfectly aligned at the same time and a balance needs to be struck in between.
	Increasing \gls{mimo} dimensions highlights the advantage of \gls{bd}-\gls{ris} for subchannel rearrangement while exposing the limitation of diagonal \gls{ris} in subspace alignment.
	This is the first paper to study \gls{bd}-\gls{ris} in general \gls{mimo} systems and discuss its shaping potential.
	% unveil and interpret the potential of \gls{bd}-\gls{ris} in general \gls{mimo} systems.

	Second, we exploit the Riemannian geometry of the Stiefel manifold and propose an efficient \gls{bd}-\gls{ris} design framework based on geodesic\footnote{A geodesic refers to the shortest path between two points in a Riemannian manifold.} \gls{rcg}.
	This method modified from \cite{Abrudan2008,Abrudan2009} not only provides better objective value and faster convergence than general non-geodesic approach \cite{Absil2009,Pan2022d}, but also works for arbitrary group size and any smooth optimization problem.
	Specifically, group-wise multiplicative rotational updates are performed along the geodesics of the Stiefel manifold and compactly evaluated as the exponential map \cite{Edelman1998}.
	By exploiting the inherent structure of unitary matrices, this method avoids retractions from the Euclidean space and facilitates the step size selection, which improves the computational efficiency and stability.
	This is the first work to tailor an efficient and universal optimization framework for \gls{bd}-\gls{ris}.

	% pioneer
	Third, we quantify the capability of a \gls{bd}-\gls{ris} to redistribute the singular values of a \gls{mimo} \gls{pc} channel.
	The {Pareto frontiers} are characterized by optimizing the {weighted sum of singular values}, where the weights can be positive, zero, or negative.
	This problem is solved by the proposed geodesic \gls{rcg} algorithm.
	The resulting singular value region generalizes most relevant metrics and provides an intuitive channel shaping benchmark.
	% We also and derive some analytical singular value bounds for rank-deficient \gls{mimo} and unitary \gls{ris}.
	To explore the shaping limits of different \gls{ris} architectures, we also derive individual and collective singular value bounds for rank-deficient channels and unitary \gls{ris} via matrix analysis.
	Results validate those bounds and show that increasing \gls{bd}-\gls{ris} group size significantly enlarges the singular value region, providing wider dynamic range and better trade-off.
	% dynamic range and trade-off of singular values are significantly improved by \gls{bd}-\gls{ris}, especially in large-scale \gls{mimo} systems.
	This is the first work to comprehensively answer the channel shaping capability question from both numerical and analytical perspectives in the most general \gls{bd}-\gls{ris} setup.
	% from a Pareto perspective.

	Fourth, we tackle \gls{bd}-\gls{ris}-aided \gls{mimo} achievable rate maximization problem with two beamforming solutions: a local-optimal approach via \gls{ao} and a low-complexity approach over channel shaping.
	The former iteratively updates active beamforming by eigenmode transmission and passive beamforming by geodesic \gls{rcg} until convergence.
	The latter suboptimally decouples the joint design into a channel power gain maximization subproblem and a conventional \gls{mimo} transmission subproblem, then propose a closed-form two-stage solution.
	We observe that \gls{bd}-\gls{ris} yields higher channel power and achievable rate than diagonal \gls{ris}, while the relative gains scale with the group size and \gls{mimo} dimensions.
	% especially with large group sizes and \gls{mimo} dimensions.
	Moreover, the rate difference between the optimal and shaping-inspired designs diminishes as the \gls{ris} evolves from diagonal to unitary.
	% Interestingly, the rate difference between the optimal and low-complexity designs diminishes as the \gls{ris} evolves from diagonal to unitary.
	Those results emphasize the importance of using \gls{bd}-\gls{ris} in large-scale \gls{mimo} systems and suggest channel shaping offers a promising path to decouple joint \gls{ris}-transceiver designs.

	Fifth, we extend the aforementioned approaches to \gls{mimo}-\gls{ic} where the \gls{bd}-\gls{ris} is used for leakage interference minimization and \gls{wsr} maximization.
	In the former case, we update the receive combiner, transmit precoder, and scattering matrix iteratively in closed forms with optimality proof.
	In the latter case, we alternatively updates the transmit precoder by bisection and the scattering matrix by \gls{rcg} algorithm.
	We also compare the impact of \gls{ris} on \gls{pc} and \gls{ic}.
	% provide a comprehensive comparison between the two approaches and discuss the implications of channel shaping in \gls{ic}.

@@

\begin{section}{\glsfmtshort{mimo}-\glsfmtshort{pc}}

	\begin{subsection}{System Model}
		Consider a \gls{bd}-\gls{ris} aided \gls{mimo} \gls{pc} system with $N_\mathrm{T}$, $N_\mathrm{S}$, $N_\mathrm{R}$ transmit, scatter, and receive antennas, respectively.
		This configuration is denoted as $N_\mathrm{T} \times N_\mathrm{S} \times N_\mathrm{R}$ in the following context.
		The \gls{bd}-\gls{ris} can be modeled as an $N_\mathrm{S}$-port network \cite{Ivrlac2010} that further divides into $G$ individual groups, each containing $L \triangleq N_\mathrm{S} / G$ elements interconnected by real-time reconfigurable components \cite{Shen2020a}.
		To simplify the analysis and explore the performance limits, we assume a lossless asymmetric network without mutual coupling between scattering elements, as previously considered in \cite{Li2023b,Li2023c,Bartoli2023}.
		The overall scattering matrix of the \gls{bd}-\gls{ris} is block-unitary
		\begin{equation}
			\mathbf{\Theta} = \mathrm{diag}(\mathbf{\Theta}_1,\ldots,\mathbf{\Theta}_G),
			\label{eq:bd_ris}
		\end{equation}
		where $\mathbf{\Theta}_g \in \mathbb{U}^{L \times L}$ is the $g$-th unitary block (i.e., $\mathbf{\Theta}_g^\mathsf{H} \mathbf{\Theta}_g = \mathbf{I}$) that describes the response of group $g \in \mathcal{G} \triangleq \{1, \ldots, G\}$.
		Note that diagonal and unitary \gls{ris} can be regarded as its extreme cases with group size $L=1$ and $L=N_\mathrm{S}$, respectively.
		Some potential physical architectures of \gls{bd}-\gls{ris} are illustrated in \cite[Fig. 3]{Shen2020a}, \cite[Fig. 5]{Li2023c}, and \cite[Fig. 2]{Nerini2024}, where the radiation pattern and circuit topology need to be modelled in the scattering matrix.

		Let $\mathbf{H}_\mathrm{D} \in \mathbb{C}^{N_\mathrm{R} \times N_\mathrm{T}}$, $\mathbf{H}_\mathrm{B} \in \mathbb{C}^{N_\mathrm{R} \times N_\mathrm{S}}$, $\mathbf{H}_\mathrm{F} \in \mathbb{C}^{N_\mathrm{S} \times N_\mathrm{T}}$ denote the direct (transmitter-receiver), backward (\gls{ris}-receiver), and forward (transmitter-\gls{ris}) channels, respectively.
		The equivalent channel is a function of the scattering matrix
		\begin{equation}
			\mathbf{H} = \mathbf{H}_\mathrm{D} + \mathbf{H}_\mathrm{B} \mathbf{\Theta} \mathbf{H}_\mathrm{F} = \mathbf{H}_\mathrm{D} + \sum_g \underbrace{\mathbf{H}_{\mathrm{B},g} \mathbf{\Theta}_g \mathbf{H}_{\mathrm{F},g}}_{\triangleq \mathbf{H}_g},
			\label{eq:channel_equivalent}
		\end{equation}
		where $\mathbf{H}_{\mathrm{B},g} \in \mathbb{C}^{N_\mathrm{R} \times L}$ and $\mathbf{H}_{\mathrm{F},g} \in \mathbb{C}^{L \times N_\mathrm{T}}$ are the backward and forward subchannels for \gls{ris} group $g$, corresponding to the $(g{-}1)L$ to $gL$ columns of $\mathbf{H}_\mathrm{B}$ and rows of $\mathbf{H}_\mathrm{F}$, respectively.
		Let $\mathbf{H}_g \triangleq \mathbf{H}_{\mathrm{B},g} \mathbf{\Theta}_g \mathbf{H}_{\mathrm{F},g}$ be the indirect channel via \gls{bd}-\gls{ris} group $g$.
		Since unitary matrices constitute an algebraic group with respect to multiplication, the scattering matrix of group $g$ can be decomposed as
		\begin{equation}
			\mathbf{\Theta}_g = \mathbf{L}_g \mathbf{R}_g^\mathsf{H},
		\end{equation}
		where $\mathbf{L}_g, \mathbf{R}_g \in \mathbb{U}^{L \times L}$ are two unitary factor matrices.
		Let $\mathbf{H}_{\mathrm{B},g} = \mathbf{U}_{\mathrm{B},g} \mathbf{\Sigma}_{\mathrm{B},g} \mathbf{V}_{\mathrm{B},g}^\mathsf{H}$ and $\mathbf{H}_{\mathrm{F},g} = \mathbf{U}_{\mathrm{F},g} \mathbf{\Sigma}_{\mathrm{F},g} \mathbf{V}_{\mathrm{F},g}^\mathsf{H}$ be the compact \gls{svd} of the backward and forward channels, respectively.
		The equivalent channel can thus be rewritten as
		\begin{equation}
			\mathbf{H} = \overbrace{\mathbf{H}_\mathrm{D} + \sum_g \mathbf{U}_{\mathrm{B},g} \mathbf{\Sigma}_{\mathrm{B},g} \underbrace{\mathbf{V}_{\mathrm{B},g}^\mathsf{H} \mathbf{L}_g \mathbf{R}_g^\mathsf{H} \mathbf{U}_{\mathrm{F},g}}_\text{backward-forward} \mathbf{\Sigma}_{\mathrm{F},g} \mathbf{V}_{\mathrm{F},g}^\mathsf{H}}^\text{direct-indirect}.
			\label{eq:channel_equivalent_svd}
		\end{equation}

		% \begin{remark}
		By analyzing \eqref{eq:channel_equivalent_svd}, we conclude that the off-diagonal entries of the \gls{bd}-\gls{ris} scattering matrix provide two key potentials for \gls{mimo} channel shaping:
		\begin{itemize}
			\item \emph{Subchannel rearrangement:} This unique feature of \gls{bd}-\gls{ris} exploits the spatial diversity by rearranging and combining the backward and forward channel branches within each group. In \gls{siso}, diagonal \gls{ris} with perfect phase matching provides a maximum indirect channel amplitude of $\sum_{n=1}^{N_\mathrm{S}} \lvert h_{\mathrm{B},n} \rvert \lvert h_{\mathrm{F},n} \rvert$ while \gls{bd}-\gls{ris} can generalize it to $\sum_{g=1}^{G} \sum_{l=1}^{L} \lvert h_{\mathrm{B},\pi_{\mathrm{B},g}(l)} \rvert \lvert h_{\mathrm{F},\pi_{\mathrm{F},g}(l)} \rvert$, where $\pi_{\mathrm{B},g}$ and $\pi_{\mathrm{F},g}$ are permutations of $\mathcal{L} \triangleq \{1, \ldots, L\}$.
			Note the first summation is over groups and the second summation is over permuted subchannels.
			By rearrangement inequality, the maximum channel gain is attained by pairing the $l$-th strongest backward and forward branches.
			% , providing up to \qty{62.5}{\percent} asymptotic power gain over diagonal \gls{ris} in \gls{siso} Rayleigh fading \cite{Shen2020a}.
			Since the number of subchannels associated with each group is proportional to $N_\mathrm{T} N_\mathrm{R}$, we conclude the advantage of \gls{bd}-\gls{ris} in subchannel rearrangement scales with \gls{mimo} dimensions,
			\item \emph{Subspace alignment:} Each group can align the singular vectors of the associated backward-forward (intra-group, multiplicative) channels and direct-indirect (inter-group, additive) channels. In \gls{siso}, subspace alignment boils down to phase matching and the optimal scattering matrix of group $g$ that maximizes the channel gain is
			\begin{equation}
				\mathbf{\Theta}_g^\star = \exp \bigl(\jmath \mathrm{arg}(h_\mathrm{D})\bigr) \mathbf{V}_{\mathrm{B},g} \mathbf{U}_{\mathrm{F},g}^\mathsf{H},
				\label{eq:scattering_siso}
			\end{equation}
			where $\mathbf{V}_{\mathrm{B},g} = \bigl[\mathbf{h}_{\mathrm{B},g}/\lVert \mathbf{h}_{\mathrm{B},g} \rVert, \mathbf{N}_{\mathrm{B},g}\bigr] \in \mathbb{U}^{L \times L}$, $\mathbf{U}_{\mathrm{F},g} = \bigl[\mathbf{h}_{\mathrm{F},g}/\lVert \mathbf{h}_{\mathrm{F},g} \rVert, \mathbf{N}_{\mathrm{F},g}\bigr] \in \mathbb{U}^{L \times L}$, and $\mathbf{N}_{\mathrm{B},g}, \mathbf{N}_{\mathrm{F},g} \in \mathbb{C}^{L \times (L-1)}$ are the orthonormal bases of the null spaces of $\mathbf{h}_{\mathrm{B},g}$ and $\mathbf{h}_{\mathrm{F},g}$, respectively.
			Diagonal \gls{ris} ($L=1$, empty null spaces) thus suffices for perfect phase matching in \gls{siso}.
			When it comes to \gls{mimo}, each individual scattering element can only apply a common phase shift to the ``pinhole'' indirect channel $\mathbf{H}_g \in \mathbb{C}^{N_\mathrm{R} \times N_\mathrm{T}}$ passing through itself.
			That is, the disadvantage of diagonal \gls{ris} in subspace alignment scales with \gls{mimo} dimensions.
			As will be shown later, even if the \gls{bd}-\gls{ris} is unitary, there still exists a tradeoff between the alignment of direct-indirect and backward-forward subspaces.
		\end{itemize}
		% \end{remark}
